---
title: "Analysis Script for School Psych Review Submission"
author: "Shaina Trevino"
date: "`r Sys.Date()`"
output: html_document
---

```{r troubleshoot, include = FALSE}
# If you are having trouble loading the environment. Open a new Rstudio (not this project) and install renv manually load the reproducible environment with the following code
#if (!require("renv")) install.packages("renv")

```

```{r renv, include = FALSE}
# Manually load the reproducible environment
# If this doesn't work the first time, try running it in the console
renv::restore()

# If you are having issues restoring renv, uncomment and run this code (or try installing Rtools)
#options(pkgType = "binary")
#renv::restore()

# If you are still having issues, you should be able to run the next code chunk to manually install and load required packages, but versions may differ

```


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, warnings = FALSE)

# Install and load pacman package (if having issues with renv)
#if (!require("pacman")) install.packages("pacman")

# Install and load required packages
pacman::p_load(tidyverse, rio, here)

# Set file path for data importing
cwd <- here::here("data")

```

## Overview

This document reproduces the quantitative results reported in our publication submitted to School Psychology Review (SPR). Qualitative results were reproduced independently by another author and are not part of this document but can be confirmed using these data files. This file is not intended to be knitted - it is designed to be run chunk by chunk to reproduce results. The first two sections are importing and data cleaning code followed by code to reproduce our quantitative results. Code/results are presented in the order in which results are presented in the manuscript with corresponding subheadings. Data were exported on 2/26/2025 for analysis. Study and review characteristics were exported directly from DistillerSR, the demographic data file was exported from DistillerSR and re-formatted in Excel. 

```{r import}
# Import demographics data collected for this manuscript
demo_df <- import(here(cwd, "2025_SPR_SchoolPsychRev_Data.xlsx"), sheet = "data")

# Import review level characteristics from DistillerSR
review_df <- import(here(cwd, "2025_SPR_review_characteristics.xlsx"), sheet = "raw_data")

# Import study level characteristics from DistillerSR
study_df <- import(here(cwd, "2025_SPR_study_characteristics.xlsx"), sheet = "raw_data")

# Import study eligibility decisions from first overview publication (per reviewer comment 5/1/25)
jccp_elig <- import(here(cwd, "2025_SPR_citation_matrix.xlsx"), sheet = "eligibility_decisions")


```

```{r tidy}
# Specify numeric variables for calculations
numeric_vars <- c("study_publication_year")

study_df <- study_df %>% 
  janitor::clean_names() %>% 
  mutate(across(all_of(numeric_vars), as.numeric),
         across(where(is.numeric), ~ na_if(., -999)))

```

## Results

Number of eligible reviews:

```{r elig-reviews}
# Extract number of reviews
nrow(review_df)

```

Number of eligible studies:

```{r elig-study}
# Extract number of included studies
nrow(study_df)

```

Range of publication years for included studies:

```{r study-year}
# Calculate min/max study publication year
study_df %>% 
  summarize(min = min(study_publication_year, na.rm = TRUE),
            max = max(study_publication_year, na.rm = TRUE))

```

Number of unique countries among included studies:

```{r study-country}
# Create df to show frequency of each country
country_counts <- study_df %>%
  pivot_longer(
    cols = starts_with("study_country"),
    names_to = "country_var",
    values_to = "country"
  ) %>%
  filter(!is.na(country) & country != "") %>%
  mutate(country = str_to_title(country)) %>%
  count(country, name = "count") %>%
  mutate(percent = round((count / sum(count)) * 100, 0)) %>%
  arrange(desc(count))

# Extract number of countries
nrow(country_counts)

```

Number of studies conducted in each country (studies with multiple countries contribute to each listed country):

```{r study-country-tbl}
# Print country summary table
country_counts

```

Number of studies reporting at least one PROGRESS domain:

```{r demo-domain}
# Specify progress domains
progress_domains <- c("place of residence", "race, ethnicity, culture, and langugage", "occupation", 
                      "gender and sex", "religion", "education", "social capital", "socioeconomic status")

# Calculate number and percent studies reporting at least one progress domain
length(unique(demo_df$refid))
paste0(round(length(unique(demo_df$refid)) / nrow(study_df) * 100, 0), "%")

```

Number of studies by number of PROGRESS domain reported:

```{r demo-results}
# Calculate number and percent of studies per number of domains reported
demo_df %>%
  group_by(refid) %>% 
  summarize(n_domains = n_distinct(domain, na.rm = TRUE)) %>% 
  ungroup() %>% 
  count(n_domains, name = "n_studies") %>% 
  mutate(percent = round(n_studies / sum(n_studies) * 100, 0))

```


### Place of Residence

Number of studies reporting place of residence:

```{r domain-place}
# Calculate number and percent of studies reporting rurality
demo_df %>% 
  filter(domain == "Place") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

### Race, Ethnicity, Culture, and Language

Number of studies reporting information on student race, ethnicity, culture, and language:

```{r domain-race}
# Calculate number and percent reporting info on race, ethnicity, culture, and language
demo_df %>% 
  filter(domain == "Race/ethnicity/culture/language") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

Number of studies reporting student language:

```{r domain-language}
# Calculate number and percent reporting student language
demo_df %>% 
  filter(domain == "Race/ethnicity/culture/language" & demographic == "Language") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

Number of studies reporting student nationality:

```{r domain-nationality}
# Calculate number and percent reporting student nationality
demo_df %>% 
  filter(domain == "Race/ethnicity/culture/language" & demographic == "Nationality") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

### Occupation

Number of studies reporting parental occupation:

```{r domain-occupation}
# Calculate number and percent of studies reporting parental occupation
demo_df %>% 
  filter(domain == "Occupation") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

### Gender

Number of studies reporting information on student gender or sex:

```{r domain-gender}
# Calculate number and percent of studies reported information on gender/sex
demo_df %>% 
  filter(domain == "Gender/sex") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

Number of studies providing options for transgender:

```{r domain-trans}
# Calculate number and percent of studies providing options for transgender
demo_df %>% 
  filter(domain == "Gender/sex",
         str_detect(measure, regex("transgender", ignore_case = TRUE))) %>% 
  summarise(unique_refid_count = n_distinct(refid))

```

Number of studies providing options for "prefer not to specify":

```{r domain-notspecified}
# Calculate number and percent of studies providing options for prefer not to specify
demo_df %>% 
  filter(domain == "Gender/sex",
         str_detect(measure, regex("prefer not to", ignore_case = TRUE))) %>% 
  summarise(unique_refid_count = n_distinct(refid))

```

Number of studies reporting student sexual orientation:

```{r domain-sexorien}
# Calculate number and percent of studies reporting sexual orientation
demo_df %>% 
  filter(domain == "Gender/sex",
         demographic == "Sexual orientation") %>% 
  summarise(unique_refid_count = n_distinct(refid))

```

### Religion

Number of studies reporting student religion:

```{r domain-religion}
# Calculate number and percent of studies reporting religion
demo_df %>% 
  filter(domain == "Religion") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

### Education

Number of studies reporting any information related to education:

```{r domain-education}
# Calculate number and percent of studies reporting education domain
demo_df %>% 
  filter(domain == "Education") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

Number of studies reporting parent education:

```{r domain-parent-edu}
# Calculate number and percent reporting parent education
demo_df %>% 
  filter(domain == "Education",
         demographic == "Parent's education") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

Number of studies reporting student special education status or needs:

```{r domain-sped}
# Calculate number and percent reporting special education
demo_df %>% 
  filter(domain == "Education",
         demographic == "Special education") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

Number of studies reporting student advanced placement status:

```{r domain-adv-plac}
# Calculate number and percent reporting advanced placement
demo_df %>% 
  filter(domain == "Education",
         demographic == "Advanced placement") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

Number of studies reporting participation in educational programs:

```{r domain-edu-prog}
# Calculate number and percent reporting school programs
demo_df %>% 
  filter(domain == "Education",
         demographic == "School program") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

### Social Capital

Number of studies reporting information on student social capital:

```{r domain-social-capital}
# Calculate number and percent of studies reporting social capital domain
demo_df %>% 
  filter(domain == "Social capital") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

Number of studies reporting each sub-domain of social capital (studies with multiple domains are counted in each category):

```{r domain-social-tbl}
# Calculate frequencies of specific social capital domains reported
demo_df %>% 
  filter(domain == "Social capital") %>% 
  group_by(demographic) %>% 
  summarize(n = n_distinct(refid),
            percent = round(n / nrow(study_df) * 100, 0))

```

### Socioeconomic Status

Number of studies reporting information on socioeconomic status (SES):

```{r domain-ses}
# Calculate number of studies reporting SES domain
demo_df %>% 
  filter(domain == "Socioeconomic status") %>% 
  summarize(n_unique_refid = n_distinct(refid),
            percent = round(n_unique_refid / nrow(study_df) * 100, 0))

```

Number of studies reporting each sub-domain of SES:

```{r domain-ses-tbl}
# Calculate frequencies of specific SES domains
demo_df %>% 
  filter(domain == "Socioeconomic status") %>% 
  group_by(demographic) %>% 
  summarize(n = n_distinct(refid),
            percent = round(n / nrow(study_df) * 100, 0))

```

## Figure 1. Heatmap of PROGRESS Domains Reported Across Included Studies

```{r heatmap-main}
# Extract refids and study names
refid_studyname <- study_df %>% 
  select(refid, study_author_year) %>% 
  rename(study = study_author_year)

# Summarize to 1 if a refid reported anything in the domain
domain_matrix <- demo_df %>%
  distinct(refid, domain) %>%
  mutate(value = 1) %>%
  pivot_wider(names_from = domain, values_from = value, values_fill = 0) %>% 
  left_join(refid_studyname) %>% 
  select(-refid) %>% 
  rename(`Race/ethnicity*` = `Race/ethnicity/culture/language`,
         SES= `Socioeconomic status`)

# Convert to long
heatmap_data <- domain_matrix %>%
  pivot_longer(-study, names_to = "domain", values_to = "reported")
heatmap_data <- heatmap_data %>%
  mutate(study = as.factor(study))


# Plot
ggplot(heatmap_data, aes(x = fct_reorder(domain, reported, .fun = sum, .desc = TRUE),
                         y = fct_rev(study), fill = factor(reported))) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c("0" = "lightgray", "1" = "steelblue"),
                    name = "Reported",
                    labels = c("No", "Yes")) +
  theme_minimal() +
  labs(x = "Domain", y = "Study") +
  theme(axis.text.y = element_text(size = 7),
        axis.text.x = element_text(size = 10))


#title = "Reporting of Demographic Domains by Study",
#caption = "*Race/Ethnicity also includes culture and language; SES = Socioeconomic status"


```

## Supplemental Materials

Supplemental materials were added per peer-reviewer feedback. 

### S1: PRISMA Flow Diagram

Unfortunately, there is not code to produce most of the numbers reported for the PRISMA flow diagram from our search since they were taken directly from DistillerSR and Zotero/FileMaker at the time of our first overview (https://doi.org/10.1037/ccp0000930). At the time, we did not have a system in place for taking snapshots of all search or screening data from DistillerSR and cannot pull those retrospectively since we have continued to update our search and screening numbers for our living review. Numbers that were pulled directly from DistillerSR include:
> Number of records identified from search: 18,083
> Duplicates removed: 3,522
> Records screened: 14,561
> Records excluded: 12,235
> Reports sought for retrieval: 2,326
> Reports not retrieved: 21
> Reports assessed for eligibility: 2,305
> Reviews and reports (k) excluded (n = 2,268, k = 2,109):
>> Anxiety focus only (k = 154)
>> Ineligible population (k = 130)
>> Ineligible interventions (k = 868)
>> Ineligible comparator (k = 55)
>> Ineligible outcomes (k = 49)
>> Ineligible setting (k = 236)
>> Ineligible study design (k = 467)
>> Ineligible publication type (k = 114)
>> Unclear – need more details (k = 36)
> Reports included: k = 37

```{r PRISMA}
# Extract number of included reviews
nrow(review_df)

# Calculate number of studies excluded
sum(jccp_elig$decision == "Exclude") + sum(jccp_elig$decision == "Unclear")

# Table of study exclude reasons (from first overview in JCCP)
summary(as.factor(jccp_elig$reason))

# Number of unclear
sum(jccp_elig$decision == "Unclear")


# Extract number of included studies
nrow(study_df)
```

### Heatmaps

Heatmaps were added per a reviewer comment to more easily summarize the patterns of reporting across all studies. 

```{r heatmap-demo-separate}
# Plot separately
# Step 1: Prepare the data
heatmap_demo <- demo_df %>%
  distinct(refid, domain, demographic) %>%
  left_join(refid_studyname) %>% 
  select(-refid) %>% 
  mutate(
    demographic = ifelse(demographic == "Free/reduced school meals", "Free/reduced meals", demographic), 
    domain = ifelse(domain == "Race/ethnicity/culture/language", "Race/ethnicity", domain), 
    refid = as.factor(study),
    demographic = as.factor(demographic),
    reported = 1
  ) %>% 
  select(-study)

# Step 2: Nest the data by domain
plots_by_domain <- heatmap_demo %>%
  # split into one tibble per domain
  group_split(domain) %>%
  set_names(map_chr(., ~ unique(.x$domain))) %>%
  map(~ {
    # 1) stash the domain name before we lose it
    domain_name <- unique(.x$domain)
    
    # 2) expand only the refids/demographics present in this slice
    df <- expand_grid(
      refid       = unique(.x$refid),
      demographic = unique(.x$demographic)
    ) %>%
      left_join(.x, by = c("refid","demographic")) %>%
      mutate(reported = replace_na(reported, 0))
    
    # 3) your ggplot, with a proper title now
    ggplot(df, aes(
             x    = fct_reorder(demographic, reported, sum, .desc = TRUE),
             y    = fct_rev(refid),
             fill = factor(reported, levels = c("0","1"))
           )) +
      geom_tile(color = "white") +
      scale_fill_manual(
        values = c("0" = "lightgray", "1" = "steelblue"),
        name   = "Reported",
        labels = c("No", "Yes"),
        drop   = FALSE
      ) +
      theme_minimal() +
      labs(
        #title = paste("Demographic Reporting for", domain_name),
        x     = "Demographic",
        y     = "Study"
      ) +
      theme(
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 9),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.title    = element_text(size = 14),  # make the “Reported” title larger
        legend.text     = element_text(size = 12),  # make the “Yes”/“No” labels larger
        legend.key.size = unit(1.2, "lines") 
      )
  })

# View and order
plots_by_domain[["Place"]]

plots_by_domain[["Race/ethnicity"]]

plots_by_domain[["Occupation"]] +
  theme(
    axis.text.y = element_text(size = 11)
  )

plots_by_domain[["Gender/sex"]] +
  theme(
    axis.text.y = element_text(size = 8)
  )

plots_by_domain[["Religion"]] +
  theme(
    axis.text.y = element_text(size = 11)
  )

plots_by_domain[["Education"]] +
  theme(
    axis.text.y = element_text(size = 11)
  )

plots_by_domain[["Social capital"]] +
  theme(
    axis.text.y = element_text(size = 11)
  )

plots_by_domain[["Socioeconomic status"]] +
  theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 11)
  )


```

